---
title: "Statistical Learning - Task 2"
author: "Sofía Gianelli"
date: "March, 2024"
output:
  html_document:
    number_sections: yes   
    toc: TRUE              
    toc_float: TRUE
subtitle: "Online Shoppers Purchasing Intention"
editor_options:
  markdown:
    wrap: 72
---

The aim of this project is to identify those individuals who will
purchased some specific product by online. In order to simplify the
work, we will assume that we are a company that sells books. We want to
investigate if is possible to have an eCommerce, so we will proceed to
analyze some variables related to the behavior of the individuals in a
website. The variables to be taken into account are:

[**Target**]{style="font-size: 20px;"}

-   **Revenue**: Indicates whether the visit resulted in revenue
    generation. When the variable is "FALSE" means that the individual
    didn't purchased at least one product, and when is equal to "TRUE"
    the individual did purchased.

[**Categorical variables**]{style="font-size: 20px;"}

-   **Month**: Represents the month of the visit.
-   **Operating Systems**: Represents the visitor's operating system.
-   **Browser**: Represents the visitor's browser.
-   **Region**: Represents the visitor's geographical region.
-   **Traffic Type**: Represents the source of traffic to the website.
-   **Visitor Type**: Represents whether the visitor is a returning or
    new visitor.
-   **Weekend**: Indicates whether the visit occurred on a weekend.

[**Numeric variables**]{style="font-size: 20px;"}

-   **Administrative**: Represents the number of administrative pages
    visited by the visitor in a session.
-   **Administrative Duration**: Represents the total time spent by the
    visitor on administrative pages.
-   **Informational**: Represents the number of informational pages
    visited by the visitor in a session.
-   **Informational Duration**: Represents the total time spent by the
    visitor on informational pages.
-   **Product Related**: Represents the number of product-related pages
    visited by the visitor in a session.
-   **Product Related Duration**: Represents the total time spent by the
    visitor on product-related pages.
-   **Bounce Rates**: Represents the percentage of visits in which the
    visitor left the site from the landing page.
-   **Exit Rates**: Represents the percentage of exits from the website.
-   **Page Values**: Represents the average value for a web page that a
    user visited before completing an e-commerce transaction.
-   **Special Day**: Indicates the closeness of the site visiting time
    to a specific special day, such as Valentine's Day or Mother's Day.

# Pre-processing

## Libraries and loading the dataset

```{r libraries, warning=FALSE,message=FALSE}
library(readxl)
library(ggplot2)
library(caret)
library(VIM)
library(knitr)
library(reshape2)
library(corrplot)
library(gridExtra)
library(pROC)
library(MASS)
library(tidyr)
library(class)
library(e1071)
library(rpart)
library(rpart.plot)
library(randomForest)
library(gbm)
set.seed(1999)
```

```{r}
dataset <- read.csv("online_shoppers_intention.csv", sep = ",")
```

```{r summary}
summary(dataset)
```

```{r string, include=FALSE}
str(dataset)
```

In this case, we observed some basic statistics about the variables. We
have 8 categorical variables and 10 numeric. In general, the variables
don't have so much variablity in their distributions.

## Duplicates and Colnames

The next step was to check if there exist duplicated rows.

```{r duplicated row}
# Checking for duplicated rows
print(paste("Duplicate rows:", sum(duplicated(dataset))))
```

As we can observe the dataset contains duplicate rows, so we will
proceed to remove it.

```{r removing duplicate rows, echo=FALSE}
# Remove duplicate rows from the dataframe
dataset <- unique(dataset)

# Check for duplicated rows in the unique dataframe
print(paste("Duplicate rows:", sum(duplicated(dataset))))
```

Now, we have 0 duplicate rows.

## Categorical variables into factor

Before spliting the dataset into train and test, we wanted to maintain
the order into the subsets so we ordered the variables like: first the
categorical variables and then the numeric ones.

```{r ordering the variables}
# Categorical first, then numeric
dataset <- dataset[,c(18,11,12,13,14,15,16,17,1,2,3,4,5,6,7,8,9,10)]
```

Before proceed to make some data visualization we had to change the
string of the categorical variables into factor form in the dataset.

```{r factorizing the variables}
dataset$Revenue <- ifelse(dataset$Revenue == "TRUE", "TRUE", "FALSE")
# Categorical as factor
dataset[, 1:8] <- lapply(dataset[, 1:8], factor)
```

## Split train and test

After cleaning the dataset, we split it into training and testing
subsets to facilitate model development and evaluation. We divided into
80% train and 20% test. As we want to do it well, we ensure that the
proportion of each class is preserved across both sets, leading to more
representative model evaluation.

```{r balance}
table(dataset$Revenue)/length(dataset$Revenue)
```

We have a 84,37% of the individuals that didn't purchased the product,
but the rest of the 15,63% purchased it.

```{r spliting data}
in_train <- createDataPartition(dataset$Revenue, p = 0.8, list = FALSE)  # 80% for training
train <- dataset[ in_train,]
test <- dataset[-in_train,]
table(train$Revenue)/length(train$Revenue)
```

After splitting the dataset using the function **createDataPartition**,
we ensured that the proportion remains exactly as before.

## Treating the NA's values

One common step of the pre-processing is to check and correct the
missing values (NA's), so the first thing that we made is corroborate if
the train and test subset had missing values.

```{r NAs in train}
sum(is.na(train))
sum(is.na(test))
```

# Exploratory Data Analysis

**Target**

```{r target plot,fig.height=4,fig.width=3, echo=FALSE}
# Calculate the percentage of each group
percentage <- prop.table(table(train$Revenue)) * 100

# Create a data frame for the labels
label_data <- data.frame(Revenue = names(percentage), 
                         Percentage = as.numeric(percentage))

ggplot(label_data, aes(x = Revenue, y = Percentage, fill = Revenue)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(Percentage, 2), "%")),
            vjust = -0.5, size = 3) +
  labs(title = "Online Shoppers", x = "Revenue", y = "Percentage",
       fill = "Count") +  
  scale_fill_manual(values = c("#D8BFD8", "#EE9572")) +  
  theme(axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        panel.background = element_rect(fill = "transparent"),
        legend.position = "none",
        plot.title = element_text(size = 8))
```

As we observed before, the are much more people who didn't purchased a
product even though they surfed in the website.

**Categorical Variables**

```{r categorical variables, echo=FALSE, warning=FALSE}
# Create a ggplot object with facet_wrap
ggplot(pivot_longer(train, cols = 2:8, names_to = "Variable"), 
       aes_string(x = "value", fill = "Revenue")) +
    geom_bar() +
  labs(x = "Value",
       y = "Frequency",
       fill = "Online Shoppers") +
  facet_wrap(~ Variable, scales = "free") +
  theme_minimal() +
  scale_fill_manual(values = c("#D8BFD8", "#EE9572"))
```

If we observe the above graphs, we could deduce that the most used
browser is the type 2, that this could be the Google Chrome browser. The
months with more views in the website are May, November and March.
However, even though the month that more people visit the website is not
the month that more people purchased the product, this month is
November. The operating system most used is the type 2 one, we will
assume that is Windows. Returning visitors are the people that visit and
purchased the most. Finally, we can observe that on weekends, the
individuals surf into the website and purchased less than in the in the
week.

In addition, we observed that the variables TraffiyType and Browser have
1 or 0 observations in one of the levels, and this can cause some
problems training the models, so we decided to remove them.

```{r}
# For TrafficType variable
train <- train[!(train$TrafficType %in% c("17", "12")), ]
train$TrafficType <- droplevels(train$TrafficType)
test <- test[!(test$TrafficType %in% c("17", "12")), ]
test$TrafficType <- droplevels(test$TrafficType)

# For Browser variable
train <- train[train$Browser != "9", ]
train$Browser <- droplevels(train$Browser)
test <- test[test$Browser != "9", ]
test$Browser <- droplevels(test$Browser)
```

**Numerical variables**

```{r correlation matrix, echo=FALSE}
numeric_train <- train[sapply(train, is.numeric)]
correlation_matrix <- cor(numeric_train)
melted_correlation <- melt(correlation_matrix)
ggplot(melted_correlation, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "#030303", mid = "#FFE4E1", high = "lightsalmon4", midpoint = 0, limit = c(-1,1),
                       space = "Lab", name="Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.justification = c(1, 0)) +
  labs(title = "Correlation Heatmap")
```

The previous heatmap is a correlation matrix of the numeric variables.
The variable "administrative" is correlated with almost every variable
except for "pagevalues". In addition, we can see that the variables that
have the number of visit and number of time spent in this type of page
are highly correlated between them. Finally, "exitpages" is negative
correlated with almost every variable, which makes sense.

The next step of this proyect is to discern the distribution of the
numeric variables to check if we have to make some transformation to
avoid problems in the future models.

```{r numeric variables, echo=FALSE}
ggplot(pivot_longer(train, cols = 9:18, names_to = "Variable"), 
       aes_string(x = "value", fill = "Revenue")) +
  geom_density(alpha = 0.8) +
  labs(x = "Value",
       y = "Density",
       fill = "Online Shoppers") +
  facet_wrap(~ Variable, scales = "free") +
  theme_minimal() +
  scale_fill_manual(values = c("#D8BFD8", "#EE9572"))
```

As we can observe, the distributions are concentrated in the left, with
a huge concentration in zero. In order to continue evaluating which
transformations make we will calculate the skewness and kurtosis of each
variable. Skewness and kurtosis are statistical measures used to
describe the shape of a distribution. Skewness measures the asymmetry of
the distribution. It tells us whether the data is skewed to the left or
right or whether it is symmetric. Kurtosis tells us whether the data is
heavy-tailed or light-tailed compared to a normal distribution.

**Skewness and Kurtosis**

```{r}
# Calculate skewness and kurtosis for each variable individually
skewness_values <- numeric(length = ncol(train[, 9:18]))
kurtosis_values <- numeric(length = ncol(train[, 9:18]))

for (i in seq_along(skewness_values)) {
  skewness_values[i] <- skewness(train[, 9:18][[i]])
  kurtosis_values[i] <- kurtosis(train[, 9:18][[i]])
}

# Create a data frame for the skewness and kurtosis values
stats_df <- data.frame(
  Variable = colnames(train[, 9:18]),
  Skewness = round(skewness_values, 2),
  Kurtosis = round(kurtosis_values, 2),  
  stringsAsFactors = FALSE
)

# Print the skewness table
kable(stats_df, caption = "Skewness and Kurtosis of Numeric Variables")
```

The skewness and kurtosis are both positive, and some variables with
high values, so we can deduce that the distributions is on the right
side with heavy tails compared to a normal distribution. After analyse
the graphs and the statistical measures we decide to transform every
numeric variable applying the logarithm + 10 to make the distribution
more symetric.

## Transformations

```{r}
# logarithmic transformations
train[,9:18] <- log(train[,9:18]+10)
```

```{r numeric variables2, echo=FALSE, warning=FALSE}
ggplot(pivot_longer(train, cols = 9:18, names_to = "Variable"), 
       aes_string(x = "value", fill = "Revenue")) +
  geom_density(alpha = 0.8) +
  labs(x = "Value",
       y = "Density",
       fill = "Online Shoppers") +
  facet_wrap(~ Variable, scales = "free") +
  theme_minimal() +
  scale_fill_manual(values = c("#D8BFD8", "#EE9572"))
```

Some variables didn't change much after applying the transformation but
some of them like ProductRelated and ProductRelated_Duration are much
better, so will continue the analysis with the variables transformed.

## Outliers

The models that we are going to perform in the following stage can be
sensitive to outliers so we tried to identify and remove them before
performing the models. To check if the variables have outliers so we
made boxplot graph to identify them.

```{r box plot, echo=FALSE}
ggplot(pivot_longer(train, cols = 9:18, names_to = "Variable"), 
       aes_string(x = "Revenue", y = "value", fill = "Revenue")) +
  geom_boxplot(alpha = 0.8) +
  labs(x = "Online Shoppers",
       y = "Value",
       fill = "Revenue") +
  facet_wrap(~ Variable, scales = "free") +
  theme_minimal() +
  scale_fill_manual(values = c("#D8BFD8", "#EE9572"))
```

As we can observe, there is some outliers, so we will use a function to
remove it.

```{r outlier function}
outlier_detection <- function(df, n, columns) {
  rows <- c()
  will_drop_train <- c()
  for (col in columns) {
    Q1 <- quantile(df[[col]], 0.25, na.rm = TRUE)
    Q3 <- quantile(df[[col]], 0.75, na.rm = TRUE)
    IQR <- Q3 - Q1
    outlier_point <- 1.5 * IQR
    rows <- c(rows, which(df[[col]] < Q1 - outlier_point | df[[col]] > Q3 + outlier_point))
  }
  row_counts <- table(rows)
  for (r in names(row_counts)) {
    if (row_counts[r] >= n) {
      will_drop_train <- c(will_drop_train, as.integer(r))
    }
  }
  return(will_drop_train)
}
```

```{r drop outliers train}
will_drop_train <- outlier_detection(train, 5, names(train)[sapply(train, is.numeric)])
head(will_drop_train, 5)
train <- train[-will_drop_train, ]
```

```{r drop outlier test}
will_drop_test <- outlier_detection(test, 5, names(test)[sapply(test, is.numeric)])
head(will_drop_test, 5)
test <- test[-will_drop_test, ]
```

We detected outliers from test and train dataset and we removed it. The
output give us the index of the outliers removed.

## Scalling

The last step before perform the model was to scale the numeric
variables, because the variables weren't in the same scale and this
might be detrimental for the performance.

```{r scalling}
train[,9:18] <- scale(train[,9:18])
test[,9:18] <- scale(test[,9:18])
```

# Modelling

## K-Nearest Neighbors

The k-Nearest Neighbors (k-NN) classifier is an algorithm that stores
the entire training dataset in memory. Then, when making predictions, it
identifies the k-nearest neighbors (data points) to the new instance
based on distance metrics, normally it use the Euclidean distance. The
majority class label among these neighbors is then assigned as the
predicted label for the new instance.

### Training

```{r ctrl}
ctrl <- trainControl(method = "cv", 
                     number = 5,
                     classProbs = F, 
                     verboseIter=F)
```

```{r knn}
knnFit <- train(Revenue ~ ., 
                method = "knn", 
                data = train,
                preProcess = c("center", "scale"),
                tuneLength = 10,
                trControl = ctrl)
```

### Predictions

```{r knn pred}
knnPred <- predict(knnFit, test)
knn_cm <- confusionMatrix(knnPred,test$Revenue)
knn_cm
```

The accuracy is 87% which is good but the number of false negative is
higher that we desire for this project.

## Neural Networks

Neural Networks is a class of machine learning models inspired by the
structure and function of the human brain. They consist of
interconnected layers of artificial neurons that process and transform
input data to produce output predictions. They're made up of layers of
"neurons" that process information, which receive input signals, apply a
mathematical transformation, and pass the result to the next layer.
These layers are typically organized into an input layer, one or more
hidden layers, and an output layer. Then the network finally makes a
prediction.

### Training

```{r nn}
# NN with 1 hidden layer
nnFit <- train(Revenue ~., 
                  method = "nnet", 
                  data = train,
                  preProcess = c("center", "scale"),
                  MaxNWts = 1000,
                  maxit = 100,
                  tuneGrid = expand.grid(size=c(2,4,6), decay=c(0.01,0.001)),
                  maximize = F,
                  trControl = ctrl)
```

### Predictions

```{r nn pred}
nnPred <- predict(nnFit, newdata=test)
nn_cm <- confusionMatrix(nnPred, test$Revenue)
nn_cm
```

The Network Neural obtained an accuracy of 87%.

## SVMs

Support Vector Machines (SVMs) are supervised learning models used, in
this case, for classification. During training, SVMs draw a line (or
plane) that separates different groups of data points. This line is
chosen so that it maximizes the space between the points of different
categories.

### Training

```{r svm}
svmFit <- train(Revenue ~., method = "svmRadial", 
                data = train,
                preProcess = c("center", "scale"),
                tuneGrid = expand.grid(C = c(.25, .5, 1),
                                      sigma = c(0.01,.05)),
                trControl = ctrl)
```

### Predictions

```{r svm pred}
svm_Pred <- predict(svmFit, newdata=test)
svm_cm <- confusionMatrix(svm_Pred, test$Revenue)
svm_cm
```

The SVM obtained an accuracy of 89%.

## Decision trees

Decision Tree is a Supervised learning technique that can be used for
solving Classification problems. It is a tree-structured classifier,
where internal nodes represent the features of a dataset, branches
represent the decision rules and each leaf node represents the outcome.
In a Decision tree, there are two nodes, which are the Decision Node and
Leaf Node. Decision nodes are used to make any decision and have
multiple branches, whereas Leaf nodes are the output of those decisions
and do not contain any further branches.

### Training

```{r dt, message = FALSE}
dt_grid <- expand.grid(.winnow = c(TRUE, FALSE),
                    .trials = c(1, 5, 7),
                    .model="tree" )

dtFit <- train(Revenue ~.,
                data=train,
                method="C5.0",
                tuneGrid = dt_grid,
                trControl = ctrl)
```

### Predictions

```{r dt pred}
DT_pred <- predict(dtFit, test)
DT_cm <- confusionMatrix(DT_pred, test$Revenue)
DT_cm
```

Even though the accuracy is 45% which is a much smaller than the other, but the error
of false negative is smaller which is good.

## Random Forests

Random Forest is a popular machine learning algorithm that belongs to
the supervised learning technique. It is based on the concept of
ensemble learning, which is a process of combining multiple classifiers
to solve a complex problem and to improve the performance of the model.
Instead of relying on one decision tree, the random forest takes the
prediction from each tree and based on the majority votes of
predictions, and it predicts the final output.

### Training

```{r rf}
rfFit <- train(Revenue ~., 
                  method = "rf", 
                  data = train,
                  preProcess = c("center", "scale"),
                  ntree = 200,
                  cutoff=c(0.7,0.3),
                  tuneGrid = expand.grid(mtry=c(6,8,10)),
                  maximize = F,
                  trControl = ctrl)
```

### Predictions

```{r rf pred}
rf_pred <- predict(rfFit, newdata=test)
rf_cm <- confusionMatrix(rf_pred, test$Revenue)
rf_cm
```

Random Forest obtained an accuracy of 89% which is a great result so
far, additionally, this model minimized the error of false negative.

## Gradient boosting

Gradient Boosting is an ensemble learning technique used for
classification. It constructs each new model to correct the errors of
the ensemble so far. It focus on the residuals, or the differences
between the actual and predicted values, of the previous models. This
process is guided by gradient descent optimization, where each new model
is trained to minimize a chosen loss function, such as mean squared
error or cross-entropy loss.

### Training

```{r gbm}
GBMFit <- gbm(ifelse(train$Revenue=="FALSE",0,1) ~., 
              data=train, 
              distribution= "bernoulli",
              n.trees=250,
              shrinkage = 0.01,
              interaction.depth=2,
              n.minobsinnode = 8)
```

### Predictions

```{r gbm pred}
threshold <- 0.5
gbm_prob <- predict(GBMFit, newdata=test, n.trees=250, type="response")
gbm_pred <- ifelse(gbm_prob > threshold, "TRUE", "FALSE")
gbm_cm <- confusionMatrix(as.factor(gbm_pred), test$Revenue)
gbm_cm
```

The accuracy of the Gradient boosting (90%) is the highest of the models
trained in this project.

# Comparative

In this part, we will compare each model performed before. We will use
the accuracy, the kappa, the roc and the auc.

## Accuracy

```{r accuracy, echo=FALSE}
# Calculate Kappa and Accuracy for each model
knn_kappa <- knn_cm$overall[2]
knn_accuracy <- knn_cm$overall[1]

nn_kappa <- nn_cm$overall[2]
nn_accuracy <- nn_cm$overall[1]

svm_kappa <- svm_cm$overall[2]
svm_accuracy <- svm_cm$overall[1]

DT_kappa <- DT_cm$overall[2]
DT_accuracy <- DT_cm$overall[1]

rf_kappa <- rf_cm$overall[2]
rf_accuracy <- rf_cm$overall[1]

gbm_kappa <- gbm_cm$overall[2]
gbm_accuracy <- gbm_cm$overall[1]

# Create a data frame with rounded values and formatted as percentages
model_results <- data.frame(
  Model = c("KNN", "Neural Network", "SVM", "Decision Tree", "Random Forest", "Gradient Boosting"),
  Kappa = c(paste0(round(knn_kappa * 100, 2), "%"),
            paste0(round(nn_kappa * 100, 2), "%"),
            paste0(round(svm_kappa * 100, 2), "%"),
            paste0(round(DT_kappa * 100, 2), "%"),
            paste0(round(rf_kappa * 100, 2), "%"),
            paste0(round(gbm_kappa * 100, 2), "%")),
  Accuracy = c(paste0(round(knn_accuracy * 100, 2), "%"),
               paste0(round(nn_accuracy * 100, 2), "%"),
               paste0(round(svm_accuracy * 100, 2), "%"),
               paste0(round(DT_accuracy * 100, 2), "%"),
               paste0(round(rf_accuracy * 100, 2), "%"),
               paste0(round(gbm_accuracy * 100, 2), "%"))
)

# Print the model results
kable(model_results)
```

Based on the provided Kappa and Accuracy values:

-   Random Forest has the highest Kappa (60%) and a relatively high
    Accuracy (88%), making it a strong contender.
-   Gradient Boosting also shows promising performance with a high
    Accuracy (90%) and a decent Kappa (53%).

## ROC

Finally, once the models were implemented, we made the ROC curves. This
is a graphical representation that illustrates the performance of a
binary classification model across various threshold settings. It plots
the true positive rate (sensitivity) against the false positive rate
(specificity) for different threshold values. The area under the ROC
curve (AUC) quantifies the model's ability to distinguish between the
two classes, with higher AUC values indicating better discrimination
performance.

```{r roc, echo=FALSE,message=FALSE}
# Compute ROC curve for each model
roc_knn <- roc(test$Revenue, as.numeric(knnPred))
roc_nn <- roc(test$Revenue, as.numeric(nnPred))
roc_svm <- roc(test$Revenue, as.numeric(svm_Pred))
roc_dt <- roc(test$Revenue, as.numeric(DT_pred))
roc_rf <- roc(test$Revenue, as.numeric(rf_pred))
roc_gbm <- roc(test$Revenue, as.numeric(gbm_prob))

# Plot ROC curves
plot(roc_knn, col="#BCEE68", main="ROC Curves for Different Models", print.thres=TRUE,lwd=3)
plot(roc_nn, col="#8B8970", print.thres=TRUE, add=TRUE,lwd=3)
plot(roc_dt, col="#458B74", print.thres=TRUE, add=TRUE,lwd=3)
plot(roc_rf, col="#CDBA96", print.thres=TRUE, add=TRUE,lwd=3)
plot(roc_gbm, col="#CD8C95", print.thres=TRUE, add=TRUE,lwd=3)
plot(roc_svm, col="#CD5C5C", print.thres=TRUE, add=TRUE,lwd=3)

# Add legend
legend("bottomright", legend=c("KNN", "NN","SVM","DT", "RF", "GBM"), col=c("#BCEE68", "#CD8C95", "#458B74", "#CDBA96", "#8B8970","#CD5C5C"), lty=1,lwd=3)
```

```{r auc, echo=FALSE}
# Compute AUC for each model
auc_knn <- auc(roc_knn)
auc_nn <- auc(roc_nn)
auc_svm <- auc(roc_svm)
auc_dt <- auc(roc_dt)
auc_rf <- auc(roc_rf)
auc_gbm <- auc(roc_gbm)

# Create a data frame with model names and AUC values
table <- data.frame(Model = c("KNN", "NN", "SVM", "DT", "RF", "GBM"),
                        AUC = round(c(auc_knn, auc_nn, auc_svm, auc_dt, auc_rf, auc_gbm),2))

# Print the table
kable(table)
```

After analyzed the models accuracy, kappa and AUC. We will continue with
Random Forest and Gradient Boosting incorporating an economic impact.

# Economic impact

Remember that this project is to investigate the possibility of having
an ecommerce, and find the individuals that will buy the product after
surfed in the website. In order to achieve this goal, we need to plan a
model focusing in reduce the false negative. Which means, reduce the
predicted "FALSE" that are indeed "TRUE", to capture the individuals
with intentions to buy the products.

Lets refresh the confusion matrix of the two model selected before.

**Random Forest:**

```{r confussion rf}
rf_cm$table
```

**Gradient Boosting**

```{r confussion gbm}
gbm_cm$table
```

To carry out this analysis, we need to assign costs or profits to
different classification outcomes. Here's a suggestion based on the
provided information:

-   Predicting a customer who will buy the product (TRUE) accurately
    yields a profit of 20%, so the cost/unit for this scenario is 0 (no
    cost or loss).
-   Predicting a customer who will not buy the product (FALSE)
    accurately incurs no profit or loss, so the cost/unit for this
    scenario is 100.
-   Predicting a customer who will buy the product inaccurately (as a
    FALSE) results in a loss of 10%, so the cost/unit for this scenario
    is 1000.
-   Predicting a customer who will not buy the product inaccurately (as
    a TRUE) leads to a loss of 200%, so the cost/unit for this scenario
    is 200.

```{r costunit}
cost.unit <- c(0, 100, 1000, 200)
```

```{r economiccost}
EconomicCost <- function(data, lev = NULL, model = NULL) {
  y.pred = data$pred 
  y.true = data$obs
  CM = confusionMatrix(y.pred, y.true)$table
  out = sum(as.vector(CM) * cost.unit) / sum(CM)
  names(out) <- c("EconomicCost")
  out
}
```

```{r economiccost2}
EconomicCost(data = data.frame(pred = rf_pred, obs = test$Revenue))
```

```{r newctrl}
newctrl <- trainControl(method = "cv", number = 5,
                     classProbs = FALSE, 
                     summaryFunction = EconomicCost,
                     verboseIter = FALSE)
```

**Random Forest**

```{r rf train2}
rf_train <- train(Revenue ~ ., 
                  method = "rf", 
                  data = train,
                  preProcess = c("center", "scale"),
                  ntree = 200,
                  cutoff = c(0.7, 0.3),
                  tuneGrid = expand.grid(mtry = c(6, 8, 10)), 
                  metric = "EconomicCost",
                  maximize = FALSE,
                  trControl = newctrl)
```

```{r rf pred2}
final_rf_pred <- predict(rf_train, newdata=test)
final_rf_cm <- confusionMatrix(final_rf_pred, test$Revenue)
final_rf_cm
```

After applying the 'EconomicCost' function, we obtained a worst accuracy
and the less predicted true that are indeed true. Moreover, we will use
this model and implement a smaller threshold to increase the 'TRUE'
predicted.

```{r new threshold}
threshold <- 0.2
final_rfProb <- predict(rf_train, newdata=test, type="prob")
final_rfPred = rep("FALSE", nrow(test))
final_rfPred[which(final_rfProb[,2] > threshold)] = "TRUE"
final_rfCM = confusionMatrix(factor(final_rfPred), test$Revenue)
final_rfCM
```

Based on this result, we observe that although the cost per unit has
increased, indicating a stricter penalty for misclassifications, the
accuracy has decreased. However, this decrease in accuracy is
accompanied by an increase in the number of individuals predicted to buy
the product, which aligns with our primary goal.

**Gradient Boosting**

```{r gbm grid, echo=FALSE}
xgb_grid = expand.grid(
  nrounds = c(500,1000),
  eta = c(0.01, 0.001), # c(0.01,0.05,0.1)
  max_depth = c(2, 4, 6),
  gamma = 1,
  colsample_bytree = c(0.2, 0.4),
  min_child_weight = c(1,5),
  subsample = 1
)
```

```{r gbm2}
final_GBMFit <- train(Revenue ~ .,  
                        data=train,
                        trControl = newctrl,
                        metric="EconomicCost",
                        maximize = F,
                        tuneGrid = xgb_grid,
                        preProcess = c("center", "scale"),
                        method = "xgbTree"
)
```

```{r gbm pred2}
threshold <- 0.5
# Get the predicted probabilities
final_gbm_prob <- predict(final_GBMFit, newdata = test, n.trees = 250, type = "prob")

# Extract the class labels with the highest probability
final_gbm_pred <- ifelse(final_gbm_prob[, "TRUE"] > threshold, "TRUE", "FALSE")

# Convert predicted and actual classes to factors with the same levels
final_gbm_pred <- factor(final_gbm_pred, levels = c("FALSE","TRUE"))
test$Revenue <- factor(test$Revenue, levels = c("FALSE","TRUE"))

# Create confusion matrix
final_gbm_cm <- confusionMatrix(final_gbm_pred, test$Revenue)
final_gbm_cm
```

The accuracy decreased compare with the accuracy of the gradient
boosting performed before, but the number of "TRUE" predicted well
increased, which was our idea.

We will change the threshold and see if we can obtain a better confusion
matrix.

```{r gbm pred3}
threshold <- 0.2
# Get the predicted probabilities
final_gbm_prob <- predict(final_GBMFit, newdata = test, n.trees = 250, type = "prob")

# Extract the class labels with the highest probability
final_gbm_pred <- ifelse(final_gbm_prob[, "TRUE"] > threshold, "TRUE", "FALSE")

# Convert predicted and actual classes to factors with the same levels
final_gbm_pred <- factor(final_gbm_pred, levels = c("FALSE","TRUE"))
test$Revenue <- factor(test$Revenue, levels = c("FALSE","TRUE"))

# Create confusion matrix
final_gbm_cm <- confusionMatrix(final_gbm_pred, test$Revenue)
final_gbm_cm
```

The number of "TRUE" predicted well increased but the accuracy is much
worst.

# Conclusion

After training and evaluating various models, we found that the Random
Forest model trained with the EconomicCost metric was the most suitable
for our objective. Despite achieving a slightly lower accuracy of 75%,
this model demonstrated superior performance in correctly predicting
individuals who actually made purchases. By setting a threshold of 0.2,
we were able to increase the number of correctly predicted "TRUE"
instances, maximizing our ability to identify potential customers who
would buy our product on the website. While the overall accuracy may
have decreased compared to other models, our focus on accurately
identifying buyers aligns with our primary goal.
